---
title: "Crime in Hungary"
author: "Tamás Nagy"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

In this project, we will explore yearly crime statistics from Hungary between 2009 and 2022.

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

# Loading the necessary packages
library(tidyverse)
library(rvest) # webscraping
library(magrittr) # for getting objects out of lists
library(broom) # tidy model outputs

# Setting the theme for plots
theme_set(theme_light())

```

# Getting the data

We are using the official statistics from the Hungarian Statistics Office. It is on a homepage, that we scrape uing the rvest package.

```{r}
crime_url <- "https://www.ksh.hu/stadat_files/iga/hu/iga0003.html"

# Get the raw data from the web
crime_raw <- 
    read_html(crime_url) |>
    # The table has a double header. This causes some trouble with variable names.
    # It is better to read all data with both headers, but not as a header but values
    html_table(header = FALSE) |> 
    # the magrittr package has this function to extract the dataframe from the list that rvest returned
    extract2(1)

# These are the original caterories in Hungarian
# I used the clipr extension of RStudio to copy the results of this following part to the clipboard
crime_raw |> 
    slice(2) |> 
    unlist() |> 
    unname()

# I pasted it to deepl to machine translate the category names. 
# I created and saved a text file that contains the English category names.

crime_header <- read_lines(here::here("eng_header.txt"))

# Now I can combine the raw data with the English header names
crime_wide <- 
    crime_raw |> 
    set_names(crime_header) |> 
    # Remove the first 2 rows (they are the headers)
    slice(-(1:2)) |> 
    # Change every variable to numeric. Note that they where characters previously.
    # I also have to remove the space first before converting to numeric.
    mutate(across(everything(), ~str_remove(., " ") |> as.numeric()))

# Create a long format dataset as well
crime_long <- 
    crime_wide |> 
    pivot_longer(-Year)

```



```{r}

crime_long |> 
    ggplot() +
    aes(x = Year, y = value, color = name) +
    geom_line(show.legend = FALSE, linewidth = 1.1) +
    # Show linear trends
    geom_smooth(method = "lm", show.legend = FALSE) +
    facet_wrap(~name, scales = "free_y") +
    scale_y_continuous(labels = scales::comma_format()) +
    labs(y = NULL, title = "Changes")
    

```

Looking at trends for individual categories

```{r}
# Create specific datasets for individual regression analyses
dui <- 
    crime_long |> 
    filter(name == "drink/driving")

total <-
    crime_long |> 
    filter(name == "Total")

drug <-
    crime_long |> 
    filter(name == "drug offencesb")

```

Fitting linear regressions for each of the three selected categories.
If we use the scale function, we will get standardized coefficients, that can be interpreted as correlations.

```{r}

lm(scale(value) ~ scale(Year), data = dui) |> 
    summary()

lm(scale(value) ~ scale(Year), data = total) |> 
    summary()

lm(scale(value) ~ scale(Year), data = drug) |> 
    summary()

```

Fitting polynomial regression on the DUI data. We use the `poly()` function to include second and third order polinomials as predictors.

```{r}
dui_model <- 
    dui |> 
    lm(value ~ poly(Year, 3), data = _)

summary(dui_model)

# The tidy function returns the coefficients as a data frame.
tidy(dui_model, conf.int = TRUE)
```

We can create predictions for the model using the `augment()` function from the `{broom}` package. if we use original data as new data, we can also include the original preditcors. 
We can visualize the model predictions based on the model predictions.

```{r}
augment(dui_model, 
        newdata = dui,
        interval = "confidence") |> 
    ggplot() +
    aes(x = Year, y = .fitted, ymin = .lower, ymax = .upper) +
    geom_ribbon(alpha = .5, fill = "cyan") +
    geom_line()

```


## Automate modeling

We can also fit models to all separate outcomes in a few lines of code. R is a functional programming language that makes this easy, through the `map()` functions (that is part of `{tidyverse}`).

```{r}
crime_nested <-
    crime_long |> 
    group_by(name) |> 
    # Nesting the dataframe by names create separate small dataframes inside the dataframe
    nest() |> 
    # We use the map function to interate through our large tibble and reach into the small tibbles inside, to fit a linear regression on each outcome. The result can be stored in a new variable inside the large tibble.
    mutate(model = map(data, ~lm(scale(value) ~ poly(Year, 3), data = .x)),
           tidy_model = map(model, tidy, conf.int = TRUE)
           ) |> 
    ungroup()

# This is how you can reach data data
crime_nested |> 
    filter(name == "criminal mischief") |> 
    unnest(data)
```

Which categories have the strongest linear, quadratic, and cubic trends?

We arrange the data by the size of the coefficient for the appropriate polynomial term to see which trends are the strongest for each category.

```{r}

crime_nested |> 
    unnest(tidy_model) |> 
    select(-data, -model) |> 
    filter(term =="poly(Year, 3)1") |> 
    arrange(estimate)

crime_nested |> 
    unnest(tidy_model) |> 
    select(-data, -model) |> 
    filter(term =="poly(Year, 3)2") |> 
    arrange(estimate)

crime_nested |> 
    unnest(tidy_model) |> 
    select(-data, -model) |> 
    filter(term =="poly(Year, 3)3") |> 
    arrange(estimate)

```


